Sequential algorithms to split and merge ultra-high resolution 3D images

### 2017-08-18 paper.tex
Integrated compression data

Clustered reads are the more direct extension of buffered slices to
blocks: they load multiple blocks in memory, concatenate them in a
buffer and write the buffer in the reconstructed image. Seeking is
reduced compared to Naive blocks since contiguous parts of the
buffer will be written without seeking. A given
block is accessed only once during the whole merging process.

@@

We used Git tag 0.1 of our \texttt{sam} library \todo{tag it}. Our
experiment scripts are available at
\url{https://github.com/big-data-lab-team/paper-sequential-split-merge/blob/master/scripts/experiment}. We
used them to split and merge using Buffered slices, Clustered reads and
Multiple reads, with 3~GB, 6~GB, 9~GB, 12~GB and 16~GB of
memory. Table~\ref{table:configs} shows the configurations of Clustered
reads and Multiple reads for each memory value, according to
Equations~\ref{eq:seeks-cluster-reads-1}
and~\ref{eq:seeks-multiple-reads}. For instance, at 3~GB, Clustered reads were in
Case 1. We also did a run with 0~GB of memory for Buffered slices and
Clustered reads, which triggered Naive slices and Naive blocks. We did 5
repetitions for each memory value. Memory values were shuffled in each
repetition, to avoid potential ordering biases such
as caching effects. To ensure equal conditions, we dropped the kernel
page, dentry and inode caches before each run (\texttt{echo 3 | sudo
  tee /proc/sys/vm/drop\_caches}). We measured the cumulative read,
write and seek time in each run, as well as the overhead time defined
as the total time minus the sum of all other times.
\begin{table}
  \centering
\begin{footnotesize}
\begin{tabular}{|c|ccccc|}
  \hline
                 & 3~GB & 6~GB & 9~GB & 12~GB & 16~GB\\
  \hline
  Clustered  reads &  1 &  2 &  2 &  2 &  3 \\
  Multiple reads &  4 &  4 &  4 &  4 &  5\\
  \hline
\end{tabular}
\end{footnotesize}
\caption{Algorithm configurations by memory values.}
\label{table:configs}
\end{table}

@@

Figure~\ref{fig:compression} shows the impact of gzip compression on
the merging time, for Multiple reads and Naive slices. As explained
before, Clustered reads could only be applied to compressed data while
in Case 3, i.e., for 16~GB, and Naive blocks could not be
used. Compression dramatically slows down merging time for all
algorithms \todo{Check that once all results are available}.

@@

% Several other factors might influence performance
I/O optimization is a holistic problem that is in practice highly
dependent on the hardware used, firmware, operating system, file
system and programming language. Caching occurs at various levels and
might always influence performance, potentially differently depending
on the split or merge algorithm used. In some disks, seek time greatly
varies with the seek distance, which would open the door to additional
opportunities for I/O optimization. Interactions between those
components might also result in performance differences particular to
a particular algorithm. To ensure the portability of our library
across systems and configurations, we focused on reducing the overall
number of seeks and ignored specific system configurations. We
demonstrated the performance of our methods on both an HDD and an SSD
disk, using state-of-the-art and widely used versions of Linux
(CentOS7) and file systems (XFS v4.5.0).

% File formats
In terms of file formats, our results demonstrate that simple imaging
formats may be split and merged without performance loss compared to
more complex formats that try to preserve spatial locality on disk,
for instance MINC 2.0 or the format based on space-filling curves
mentioned in~\cite{burns2013open}. This is of major interest in the
current open-science context since simpler formats favor data-sharing
and interoperability.  Moreover, our algorithms could potentially be
adapted to any split geometry, even though we demonstrated them on
slices and blocks only, while file formats inevitably assume a
particular geometry. For instance, the format in~\cite{burns2013open}
is not designed to naively split slices. However, we aimed to extract
all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
single block from a large image. MINC 2.0 might also help with
on-the-fly compression for Clustered reads.

>>>>

Clustered reads are the more direct extension of Buffered slices to
blocks: they load multiple blocks in memory, concatenate them in a
buffer and write the buffer in the reconstructed image. Seeking is
reduced compared to Naive blocks since contiguous parts of the
buffer will be written without seeking. A given
block is accessed only once during the whole merging process.

@@

We used Git tag 0.1 of our \texttt{sam} library \todo{tag it}. Our
experiment scripts are available at
\url{https://github.com/big-data-lab-team/paper-sequential-split-merge/blob/master/scripts/experiment}. We
used them to split and merge using Buffered slices, Clustered reads and
Multiple reads, with 3~GB, 6~GB, 9~GB, 12~GB and 16~GB of
memory. Table~\ref{table:configs} shows the configurations of Clustered
reads and Multiple reads for each memory value, according to
Equations~\ref{eq:seeks-cluster-reads-1}
and~\ref{eq:seeks-multiple-reads}. For instance, at 3~GB, Clustered reads were in
Case 1. We also did a run with 0~GB of memory for Buffered slices and
Clustered reads, which triggered Naive slices and Naive blocks. We did 5
repetitions for each memory value. Memory values were shuffled in each
repetition, to avoid potential ordering biases such
as caching effects. To ensure equal conditions, we dropped the kernel
page, dentry and inode caches before each run (\texttt{echo 3 | sudo
  tee /proc/sys/vm/drop\_caches}). We measured the cumulative read,
write and seek time in each run, as well as the overhead time defined
as the total time minus the sum of all other times.

Compressed blocks and slices were also merged into a compressed
image. On-the-fly gzip compression was used for Naive slices, Buffered
slices and Multiple reads. As explained before, Clustered reads could
only be applied to compressed data while in Case 3, i.e., for 16~GB,
and Naive blocks could not be used at all with on-the-fly
compression. To give a reference, with used Naive blocks with
\emph{offline} compression, that it, we wrote an uncompressed image
and compressed it sequentially afterwards. \todo{confirm that.}

\begin{table}
  \centering
\begin{footnotesize}
\begin{tabular}{|c|ccccc|}
  \hline
                 & 3~GB & 6~GB & 9~GB & 12~GB & 16~GB\\
  \hline
  Clustered  reads &  1 &  2 &  2 &  2 &  3 \\
  Multiple reads &  4 &  4 &  4 &  4 &  5\\
  \hline
\end{tabular}
\end{footnotesize}
\caption{Algorithm configurations by memory values.}
\label{table:configs}
\end{table}

@@

Figure~\ref{fig:compression} shows the impact of gzip compression on
the merging time, for Multiple reads and Naive slices.
Unsurprisingly, compression dramatically slows down merging time for
all algorithms. Multiple reads, however, still provide important
speed-up compared to Naive blocks \todo{check that}, although they do
not completely reach the performance of Buffered slices and Naive
slices.

@@

% Several other factors might influence performance
I/O optimization is a holistic problem that is in practice highly
dependent on the hardware used, firmware, operating system, file
system and programming language. Caching occurs at various levels and
might always influence performance, potentially differently depending
on the split or merge algorithm used. In some disks, seek time greatly
varies with the seek distance, which would open the door to additional
opportunities for I/O optimization. Interactions between those
components might also result in performance differences particular to
a specific algorithm. To ensure the portability of our library
across systems and configurations, we focused on reducing the overall
number of seeks and ignored specific system configurations. We
demonstrated the performance of our methods on both an HDD and an SSD
disk, using state-of-the-art and widely used versions of Linux
(CentOS7) and file systems (XFS v4.5.0).

% File formats
In terms of file formats, our results demonstrate that simple imaging
formats may be split and merged without performance loss compared to
more complex formats that can preserve spatial locality on disk,
for instance MINC 2.0 or the format based on space-filling curves
mentioned in~\cite{burns2013open}. This is of major interest in the
current open-science context since simpler formats favor data-sharing
and interoperability.  Moreover, our algorithms could potentially be
adapted to any split geometry, even though we demonstrated them on
slices and blocks only, while file formats inevitably assume a
particular geometry. For instance, the format in~\cite{burns2013open}
is not designed to naively split slices. However, we aimed to extract
all the blocks, whereas, ~\cite{burns2013open} aimed to extract a
single block from a large image. MINC 2.0 might also help with
on-the-fly compression for Clustered reads.
